Exploratory Plots

### 2020-10-05 exploratory/2020/sub3_fit_subsets_fixed.Rmd
Plot de novo results alongside optifit

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE,
                      fig.path = "figures/"
                      )
```
```{r lib}
library(cowplot)
library(ggtext)
library(here)
library(knitr)
library(tidyverse)
theme_set(theme_classic())
color_palette <- RColorBrewer::brewer.pal(4, "Dark2")
dataset_colors <- c(
  human = color_palette[[3]],
  marine = color_palette[[1]],
  mouse = color_palette[[4]],
  soil = color_palette[[2]]
)
```
```{r download_data}
dat <- c('https://raw.githubusercontent.com/SchlossLab/OptiFitAnalysis/4fa7127bb6bacaa63657f0c18c0a8c6b6d18ebe7/subworkflows/3_fit_sample_subset/results/sensspec.tsv?token=AEHR6TOTYRF3XTCACDHGFRK7QRYH2',
              'https://raw.githubusercontent.com/SchlossLab/OptiFitAnalysis/4fa7127bb6bacaa63657f0c18c0a8c6b6d18ebe7/subworkflows/3_fit_sample_subset/results/benchmarks.tsv?token=AEHR6TLJWOKI4YC6OJ6BRMC7QRYXY',
              'https://raw.githubusercontent.com/SchlossLab/OptiFitAnalysis/4fa7127bb6bacaa63657f0c18c0a8c6b6d18ebe7/subworkflows/3_fit_sample_subset/results/input_sizes.tsv?token=AEHR6TJVM5FBTZRRY4USOYS7QRY6U',
              'https://raw.githubusercontent.com/SchlossLab/OptiFitAnalysis/4fa7127bb6bacaa63657f0c18c0a8c6b6d18ebe7/subworkflows/3_fit_sample_subset/results/gap_counts.tsv?token=AEHR6TMQKLHG2RNGUZQI35C7QRY5G',
              'https://raw.githubusercontent.com/SchlossLab/OptiFitAnalysis/4fa7127bb6bacaa63657f0c18c0a8c6b6d18ebe7/subworkflows/3_fit_sample_subset/results/fraction_reads_mapped.tsv?token=AEHR6TIY56KPVNGRKCOAYI27QRZBW') %>% 
  lapply(read_tsv) %>% 
  reduce(full_join) %>% 
  mutate(num_total_seqs = num_ref_seqs + num_sample_seqs,
         gaps_frac = n_gaps / total_chars)
```

@@

```{r fit_ref_frac}
dat %>%
  ggplot(aes(x = ref_frac, y = mcc, color = dataset)) +
  geom_jitter(size = 1, alpha = 0.3, width = 0.01) +
  stat_summary(fun = mean, geom = 'crossbar') +
  scale_color_manual(values = dataset_colors) +
  facet_grid(method ~ ref_weight) +
  labs(title='OptiFit Performance',
       x='reference fraction')
```

@@

```{r runtime_ref_seqs}
dat %>% ggplot(aes(x=num_ref_seqs, y=s, color=dataset)) +
  geom_point(alpha = 0.3) +
  facet_grid(method ~ ref_weight) +
  scale_color_manual(values = dataset_colors) +
  scale_x_continuous(breaks = seq(0, 200000, 50000),
                     labels = c('0', '50k', '100k', '150k', '200k')) +
  labs(title = 'Runtime over reference size',
       x = '# sequences in reference',
       y = 'seconds')
```

@@

## Fraction of sequences that map to the reference
```{r fraction_reads_mapped}
dat %>% 
  filter(method == 'closed') %>% 
  group_by(dataset, ref_weight, ref_frac) %>% 
  ggplot(aes(x=ref_frac, y=fraction_mapped, color=dataset)) +
  geom_jitter(alpha = 0.5, width = 0.01, size=1) +
  scale_color_manual(values = dataset_colors) +
  facet_wrap("ref_weight") +
  ylim(0, 1) +
  labs(title="Sequences mapped during closed-reference OptiFit",
       x='reference fraction',
       y='fraction mapped')
```

>>>>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE,
                      fig.path = "figures/"
                      )
```
```{r lib}
library(cowplot)
library(ggtext)
library(here)
library(knitr)
library(tidyverse)
theme_set(theme_classic())
color_palette <- RColorBrewer::brewer.pal(4, "Dark2")
dataset_colors <- c(
  human = color_palette[[3]],
  marine = color_palette[[1]],
  mouse = color_palette[[4]],
  soil = color_palette[[2]]
)
full_join_files <- function(filenames) {
  filenames %>% 
  lapply(read_tsv) %>% 
  reduce(full_join)
}
```
```{r download_data}
dat <- c('https://raw.githubusercontent.com/SchlossLab/OptiFitAnalysis/4fa7127bb6bacaa63657f0c18c0a8c6b6d18ebe7/subworkflows/3_fit_sample_subset/results/sensspec.tsv?token=AEHR6TOTYRF3XTCACDHGFRK7QRYH2',
              'https://raw.githubusercontent.com/SchlossLab/OptiFitAnalysis/4fa7127bb6bacaa63657f0c18c0a8c6b6d18ebe7/subworkflows/3_fit_sample_subset/results/benchmarks.tsv?token=AEHR6TLJWOKI4YC6OJ6BRMC7QRYXY',
              'https://raw.githubusercontent.com/SchlossLab/OptiFitAnalysis/4fa7127bb6bacaa63657f0c18c0a8c6b6d18ebe7/subworkflows/3_fit_sample_subset/results/input_sizes.tsv?token=AEHR6TJVM5FBTZRRY4USOYS7QRY6U',
              'https://raw.githubusercontent.com/SchlossLab/OptiFitAnalysis/4fa7127bb6bacaa63657f0c18c0a8c6b6d18ebe7/subworkflows/3_fit_sample_subset/results/gap_counts.tsv?token=AEHR6TMQKLHG2RNGUZQI35C7QRY5G',
              'https://raw.githubusercontent.com/SchlossLab/OptiFitAnalysis/4fa7127bb6bacaa63657f0c18c0a8c6b6d18ebe7/subworkflows/3_fit_sample_subset/results/fraction_reads_mapped.tsv?token=AEHR6TIY56KPVNGRKCOAYI27QRZBW') %>% 
  full_join_files() %>% 
  mutate(num_total_seqs = num_ref_seqs + num_sample_seqs,
         gaps_frac = n_gaps / total_chars)

denovo <- c('https://raw.githubusercontent.com/SchlossLab/OptiFitAnalysis/4fa7127bb6bacaa63657f0c18c0a8c6b6d18ebe7/subworkflows/1_prep_samples/results/sensspec.tsv?token=AEHR6TOXMAB75QDAO7CGTA27QR4KQ',
            'https://raw.githubusercontent.com/SchlossLab/OptiFitAnalysis/4fa7127bb6bacaa63657f0c18c0a8c6b6d18ebe7/subworkflows/1_prep_samples/results/benchmarks.tsv?token=AEHR6TI35D2TNWARGGL37W27QR4MM',
            'https://raw.githubusercontent.com/SchlossLab/OptiFitAnalysis/4fa7127bb6bacaa63657f0c18c0a8c6b6d18ebe7/subworkflows/1_prep_samples/results/dataset_sizes.tsv?token=AEHR6TJKE2C7SRDUZSGRTCC7QR4MG') %>% 
  full_join_files() %>% 
  mutate(ref_frac = 1)

denovo_sum <- denovo %>% group_by(dataset) %>% summarize(mean_mcc=mean(mcc), 
                                                         sd_mcc=sd(mcc),
                                                         mean_sec=mean(s),
                                                         sd_sec=sd(s),
                                                         num_ref_seqs = mean(num_seqs))
denovo_sum
```

@@

```{r fit_ref_frac}
dat %>%
  ggplot(aes(x = ref_frac, y = mcc, color = dataset)) +
  geom_jitter(size = 1, alpha = 0.3, width = 0.01) +
  stat_summary(fun = mean, geom = 'crossbar') +
  scale_color_manual(values = dataset_colors) +
  facet_grid(method ~ ref_weight) +
  labs(title='OptiFit Performance',
       x='reference fraction')
dat %>%
  ggplot(aes(x = ref_frac, y = mcc, color = dataset)) +
  geom_jitter(size = 1, alpha = 0.3, width = 0.01) +
  stat_summary(fun = mean, geom = 'crossbar') +
  geom_hline(aes(yintercept = mean_mcc, color = dataset), denovo_sum) +
  scale_color_manual(values = dataset_colors) +
  facet_grid(method ~ ref_weight) +
  labs(title='OptiFit Performance',
       x='reference fraction',
       caption='Sample fraction = 0.2 for OptiFit runs. The horizontal line represents _de novo_ clustering for the whole dataset.') +
  theme(plot.caption = element_markdown())
dat %>%
  ggplot(aes(x = ref_frac, y = mcc, color = ref_weight)) +
  geom_jitter(size = 1, alpha = 0.3, width = 0.01) +
  stat_summary(fun = mean, geom = 'crossbar') +
  geom_hline(aes(yintercept = mean_mcc), denovo_sum) +
  facet_grid(dataset ~ method) +
  labs(title='OptiFit Performance',
       x='reference fraction',
       caption='Sample fraction = 0.2 for OptiFit runs. The horizontal line represents _de novo_ clustering for the whole dataset.') +
  theme(plot.caption = element_markdown())
```

@@

```{r runtime_ref_seqs}
dat %>% ggplot(aes(x=num_ref_seqs, y=s, color=dataset)) +
  geom_point(alpha = 0.3) +
  facet_grid(method ~ ref_weight) +
  scale_color_manual(values = dataset_colors) +
  scale_x_continuous(breaks = seq(0, 200000, 50000),
                     labels = c('0', '50k', '100k', '150k', '200k')) +
  labs(title = 'Runtime over reference size',
       x = '# sequences in reference',
       y = 'seconds')
dat %>% group_by(num_ref_seqs, dataset, ref_weight, method) %>% 
  summarize(mean_sec=mean(s)) %>% 
  ggplot(aes(x=num_ref_seqs, y=mean_sec, color=ref_weight)) +
  geom_point(alpha = 0.5) +
  geom_point(aes(x=num_ref_seqs, y=mean_sec, color=NULL), denovo_sum) +
  facet_grid(dataset ~ method, scales = 'free') +
  #scale_color_manual(values = dataset_colors) +
  scale_x_continuous(breaks = seq(0, 200000, 50000),
                     labels = c('0', '50k', '100k', '150k', '200k')) +
  labs(title = 'Runtime over reference size',
       x = '# sequences in reference',
       y = 'seconds',
       caption = 'Sample fraction = 0.2 for OptiFit runs. The black dot represents _de novo_ clustering for the whole dataset') +
  theme(plot.caption = element_markdown())
```

@@

## Fraction of sequences that map to the reference
```{r fraction_reads_mapped}
dat %>% 
  filter(method == 'closed') %>% 
  group_by(dataset, ref_weight, ref_frac) %>% 
  ggplot(aes(x=ref_frac, y=fraction_mapped, color=dataset)) +
  geom_jitter(alpha = 0.5, width = 0.01, size=1) +
  scale_color_manual(values = dataset_colors) +
  facet_wrap("ref_weight") +
  ylim(0, 1) +
  labs(title="Sequences mapped during closed-reference OptiFit",
       x='reference fraction',
       y='fraction mapped')
dat %>% 
  filter(method == 'closed') %>% 
  group_by(dataset, ref_weight, ref_frac) %>% 
  ggplot(aes(x=ref_frac, y=fraction_mapped, color=ref_weight)) +
  geom_jitter(alpha = 0.5, width = 0.01, size=1) +
  #scale_color_manual(values = dataset_colors) +
  facet_wrap("dataset", nrow=1) +
  ylim(0, 1) +
  labs(title="Sequences mapped during closed-reference OptiFit",
       x='reference fraction',
       y='fraction mapped')
```
